Bayes Computational Methods

### 2020-03-04 Paper.Rmd
adding outlines

>>>>

# Introduction

## The Bayes Paradigm

In Bayesian statistical inference, we are interested in finding the posterior distribution of a certain parameter, given a prior distribution and a likelihood for that parameter from a sample of data. Bayes Rule states that, given a parameter of interest $\theta$ and some known data $x$, we can find the conditional probability of $\theta$ given $y$ using $$ p(\theta | x) = \frac{p(x | \theta) p(\theta)}{p(x)},$$ where in the discrete case $p(x) = \sum_\theta p(\theta) p(x|\theta),$ and in the continuous case, $p(x) = \int p(\theta) p(y|\theta) d\theta$. More formally, we can derive the posterior distribution of a parameter $\theta$, $p(\theta | x)$, by $$p(\theta | x_1, x_2, \dots, x_n) = \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}},$$ where $\ell(x | \theta)$ is proportional to the likelihood of $\theta$, conditioned on the data, and $\pi(\theta)$ is the prior distribution of $\theta$. For example, suppose that $X_i \sim \text{Bern}(\theta)$, and we are interested in inference on $\theta$. A reasonable prior to assign to $\theta$ would be a $\text{Beta}(\alpha, \beta)$, distribution, given that $\theta$ is constrained to be between 0 and 1, which is a property of the Beta distribution. Therefore, given that $\theta \sim \text{Beta}(\alpha, \beta)$ for some known values of $\alpha$ and $\beta$, then 
\begin{align*}
  p(\theta | x) &= \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}} \\
  &= \frac{\theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1}}{\int \theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1} d \theta} \\
  &\propto \theta ^ {\sum (x_i) + \alpha - 1} (1 - \theta) ^ {n - \sum (x_i) + \beta - 1},
\end{align*}
which is the kernel of a $\text{Beta}\left(\sum (x_i) + \alpha, n - \sum (x_i) + \beta \right)$ distribution. Therefore, we can analytically calculate the posterior distribution of $\theta$ as $$p(\theta | x) \sim \text{Beta}\left(\sum (x_i) + \alpha, n - \sum (x_i) + \beta\right).$$

In more complicated scenarios, it is not easy to calculate the posterior distribution of a parameter by hand, and in many cases it is impossible. This is where computing can help us.
